---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---


##  Note: Please run the source file first and any additional library
```{r setup, message = F }
knitr::opts_chunk$set(echo = T, message = F) #show code but no messages
```

```{r}
source('../UI paper 1 source.R')
library(MatchIt, quietly = T)
library(Matching, quietly = T)
library(stargazer, quietly = T)
library(lme4)
```

These are the graphing libraries and routines

```{r}
pkgs <- c('tidyverse', 'ggplot2', 'gridExtra')
lapply(pkgs, require, character.only = T)

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

```



##  1) The sample

We first filter the data to non-isle and non-crossborder TTWAs. Then we only keep
TTWA with pops over 70k in 2004.

```{r}
ttwa.tab <- '../Results/Duncan index by TTWA.csv' %>% read.csv(stringsAsFactors = F)
crossborder.ttwa <- c('Hawick', 'Kelso & Jedburgh', 'Berwick', 'Carlisle') #cross national ttwas to be omitted
isles.ttwa <- c(
  'Eilean Siar',
  'Mull & Islay',
  'Campbeltown',
  'Penzance & Isles of Scilly',
  'Orkney Islands',
  'Shetland Islands',
  'Skye & Lochalsh',
  'Lochaber',
  'Irvine & Arran',
  'Isle of Wight')
#islands


ttwa.tab <- 
  ttwa.tab %>% 
  filter(!(ttwa %in% crossborder.ttwa)) %>% #not a cross border ttwa
  filter(!(ttwa %in% isles.ttwa)) %>% #nor isle
  ##  We need to take these out first since crossborder ttwa are duplicated
  group_by(ttwa) %>%
  mutate(
    pop04 = total.pop[year == '2004']) %>% #variable for pop in 2004
  mutate(dist_nearest04 = dist_nearest[year == '2004']) %>%
  mutate(di04 = di[year == '2004']) %>%
  ungroup %>%
  filter(pop04 > 70000) %>% ## over 70k in 2004
  filter(!is.na(ttwa)) ##finally no missign ttwa as this code is just due to our previous routine

```

We can restrict our data to 2004 to compare pre- and post- matching results.
```{r}
unmatch.tab <- 
  ttwa.tab %>% 
  filter(year == '2004')

unmatch.country <-
  unmatch.tab$country %>% table

unmatch.country %>% print

```

After omitting we get `r unmatch.country[1]` English TTWAs and `r unmatch.country[2]` Scottish TTWAs.



##  2) Matching using Genmatch

We have the full sample and now we further restrict to matched data. We match on 
2004 / time invariant. We can automatically try to ger the best match using genetic matching. Then we 
can compare the stae of our matches. Genetic matchign will use the propensity score 
in the matching by default (it generate own: i think from Sekhon's examples).

```{r, results= 'hide'}
## We need to hide all the print statement caused by genmatch
match.form <- I(country == 'Scotland') ~ total.pop + total.area + dist_nearest04 + di04

set.seed(123) ## genetic matchhing is a random process so we must set seed
match.res <- 
  matchit(match.form,
          ratio = 2,
          data = unmatch.tab %>% dplyr::select(ttwa:total.area, dist_nearest04, di04),
          method = 'genetic')


          
match.stats <-
  match.res %>% summary

```

Now we can try to summarise the matching results. 

```{r}
match.stats %>% print

```

We have a total of `r match.stats$nn[1, 1]` English and `r match.stats$nn[1, 2]` Scottish ttwa. In our final matched sample, we have `r match.stats$nn[2, 1]` English TTWAs matched to`r match.stats$nn[2, 2]` Scottish ttwas. 

##  Comparing variance between indicies.

We note that the various indices are related to each other. 

```{r, message = T}
cor.df <- 
  unmatch.tab %>% filter(year == 2004) 

##  Making the correlation table 
cor.vars <- c('dist_nearest', 'di','live.in', 'crime', 'geo', 'pm25', 'work')
cor.table <- cor(cor.df %>% dplyr::select(cor.vars), 
                 method = 'spearman', 
                 use = 'pairwise.complete.obs') # only modest correlation

##  Formatting it for stargazer
cor.table[cor.table %>% upper.tri(diag = T)] <- NA
ind.nms <- c('RCI', 'Diss. Index', 'Housing', 'Crime', 'Geo. access', 'Air pollution', 'Work access')
rownames(cor.table) <- ind.nms
colnames(cor.table) <- ind.nms

stargazer(cor.table, 
          type = 'text', #anything else stargazer will output like code
          title = 'Spearman\'s rank correlation matrix between indicies (2004)',
          out = '../Results/ Index correlation table.html',
          notes = 'Note: Scottish crime data excluded from calculation')

```


##  Graphing trends over time
# Plots and generic plotting functions

```{r}
##  We want to summarise the table stats before passign to ggplot

t.val = 1.96 * sqrt(2) / 2 # Set level of the tvalue

summarise_mydata <-
  function(summariseThis = NULL, ## This is the variable to summarise by
           data = ttwa.tab
  ){
    
  ## Goal: Take in data and a name for the variable we want to summarise (called summarise this)
  ##  and end up with a summary table by county and year for:
  ##  mean of summariseThis
  ##  error bars that reflect the interyear variance and SE of TTWAs (so we have
  ##  partition errors var(y|ttwa)
  
  tempData <- data
  tempData$tempY <- data[[summariseThis]]
    
  form <- summariseThis %>% paste('~ ttwa') %>% as.formula
  
  
  residualSD <-
      lm(form, tempData) %>% residuals %>% sd 

  summarisedForGGplot <-
  tempData %>% 
  group_by(country, year) %>%
  summarise(n = length(tempY),
            y = mean(tempY),
            ymin = y - t.val * (residualSD/ sqrt(n)), 
            ymax = y + t.val * (residualSD/ sqrt(n)) 
  )
  
  return(summarisedForGGplot)
  }  

```

##  Trends in segregation


```{r}

rci.gg <- 
  ggplot(data = summarise_mydata('dist_nearest'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +
  ylab('Centralisation (RCI)') 

di.gg <- 
  ggplot(data = summarise_mydata('di'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +
  ylab('Uneveness (DI)')

## One big plot
mylegend<-g_legend(rci.gg + theme(legend.direction = 'horizontal'))

seg.grid <- grid.arrange(rci.gg + xlab('') + theme(legend.position="none"), 
                         di.gg + xlab('')+ theme(legend.position="none"), 
                         ncol = 2,
                         #                    top = 'Changes in Centralisation and Uneveness',
                         left = 'Index',
                         bottom = 'Year'
)

grid.arrange(mylegend, seg.grid, nrow=2,heights=c(1, 12),
             top = 'Change in segregation (Error bars for annual comparison)')


```


##  Inequality indices

```{r}

house.gg <- 
  ggplot(data = summarise_mydata('live.in'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +
  ylab('Housing') 


crime.gg <- 
  ggplot(data = summarise_mydata('crime'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +
  ylab('Crime')

##
geo.gg <- 
  ggplot(data = summarise_mydata('geo'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +
  ylab('Geo. access')

air.gg <- 
  ggplot(data = summarise_mydata('pm25'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +
  ylab('Pollution') 

work.gg <- 
  ggplot(data = summarise_mydata('work'), 
         mapping =aes(x = year, y = y, colour = country)) + 
  geom_errorbar(aes(ymax = ymax, ymin = ymin), alpha = 0.5) +
  geom_line(size = 1) +

  ylab('Work access') +
  xlab('')# 

##  Arrange ggplots

mylegend<-g_legend(house.gg + theme(legend.direction = 'horizontal'))


gfx <- grid.arrange(house.gg + xlab('') + theme(legend.position="none"), 
                    crime.gg + xlab('') + theme(legend.position="none"), 
                    geo.gg + xlab('')+ theme(legend.position="none"), 
                    air.gg + xlab('') + theme(legend.position="none"), 
                    work.gg + xlab('') + theme(legend.position="none"), 
                    left = 'Spatial inequality index',
                    bottom = 'Year'
)

grid.arrange(mylegend, gfx, nrow=2,heights=c(1, 12),
             top = 'Change in spatial inequality (Error bars for annual comparison)')


```



##  Regression models

The data we are using for the models

```{r}
reg.unmatch <- ttwa.tab # just the whole data

##  For the matched data we basically need the TTWA and their weights
temp.match <- MatchIt::match.data(match.res)
head(temp.match)
temp.match <- temp.match %>% dplyr::select(ttwa, weights)
reg.match <- ttwa.tab %>% merge(temp.match, by = 'ttwa')


reg.match <-
  reg.match %>% filter(year %in% c('2015', '2016'))

##  We need to use weights in our regression routine for matched data so we can only access the latest years

```



```{r}
##  The multilevel models 
form1 <- ~ I(year - 2004) + I(year - 2004)*country + (1|ttwa)
form2 <- ~ country + scale(total.pop) + scale(total.area) + dist_nearest04 + di04 ## We omit the I(year - 2004) variable since last 
##  years 2015 or 2016 is associated with country
##Note:  form2 is form1 with more variables

# Segregation indicies regression ------------------------------------------------
##  
dist.reg <- lmer(form1 %>% update(dist_nearest ~ .), reg.unmatch)
dist.reg2 <- lm(form2 %>% update(dist_nearest ~ .), reg.match, weights = weights)

di.reg <- lmer(form1 %>% update(di ~ .), reg.unmatch)
di.reg2 <- lm(form2 %>% update(di ~ .), reg.match, weights = weights)


## Save table
stargazer(dist.reg,
          di.reg, 
          type = 'text',
          title = 'Model results for segregation (unmatched random effects)',
          dep.var.caption = 'Segregation index',
          dep.var.labels.include = F,
          covariate.labels = c('Year (2004 = 0)', 'Scotland', 'Year:Scotland'),
          column.labels = c('RCI', 'DI'),
          omit = c('scale', 'dist_nearest04', 'di04'),
          no.space = T,
          out = '../Results/ Segregation index model results unmatched.html',
          keep.stat = c('n')
)

stargazer(dist.reg2,
          di.reg2, 
          type = 'text',
          title = 'Model results for segregation (matched OLS)',
          dep.var.caption = 'Segregation index',
          dep.var.labels.include = F,
          covariate.labels = c('Scotland', 'Constant'),
          column.labels = c('RCI', 'DI'),
          omit = c('scale', 'dist_nearest04', 'di04'),
          no.space = T,
          out = '../Results/ Segregation index model results matched.html',
          keep.stat = c('n')
)


```

##  Comparison between segregation by social housing

```{r}
ttwa.tab_HH <- '../Results/Duncan index by TTWA Households.csv' %>% read.csv(stringsAsFactors = F)
compare.tab <-
  ttwa.tab %>%
  left_join(ttwa.tab_HH,
            by = c('ttwa', 'country', 'year'),
            suffix = c('.inc', '.HH')
  )


compare.tab %>% 
  group_by(country, year) %>%
  summarise(rci_HH =  dist_nearest.HH %>% mean,
            di.HH = di.HH %>% mean)


##  Correlation between social housing and decentralisaiton at 2004
## Data prep

summary.tab <-
  compare.tab %>% 
  group_by(ttwa) %>%
  summarise(
    change_rci.inc = dist_nearest.inc[year >= 2015] - dist_nearest.inc[year == 2004],
    change_rci.HH = dist_nearest.HH[year >= 2015] - dist_nearest.HH[year == 2004],

    change_di.inc = di.inc[year >= 2015] - di.inc[year == 2004],
    change_di.HH = di.HH[year >= 2015] - di.HH[year == 2004]
    
  )

##  Over time correlation
summary.tab[-1] %>%  cor(use= "pairwise.complete.obs", method = 'spearman') #mega low 

##  Correlation at 2004
compare.tab %>% 
  filter(year == '2004') %>% 
  dplyr::select(starts_with('dist_nearest.'), starts_with('di.')) %>%
  cor(use= "pairwise.complete.obs", method = 'spearman')


```




##  Regression by index

```{r}

in.reg <- lmer(form1 %>% update(live.in ~ .), reg.unmatch)

crime.reg <- lmer(form1 %>% update(crime ~ .), reg.unmatch)

geo.reg <- lmer(form1 %>% update(geo ~ .), reg.unmatch)

pm25.reg <- lmer(form1 %>% update(pm25 ~ .), reg.unmatch)

work.reg <- lmer(form1 %>% update(work ~ .), reg.unmatch)

reg.list <- list(in.reg, 
                 crime.reg, 
                 geo.reg, 
                 pm25.reg,
                 work.reg)

stargazer(reg.list, type = 'text',
          title = 'Random effect model results',
          dep.var.caption = 'Inequality index',
          dep.var.labels = c('Housing', 'Crime', 'Geo. access', 'Air pollution', 'Work access'),
          covariate.labels = c('Year (2004 = 0)', 'Scotland', 'Year:Scotland'),
#          column.labels = rep(c('All', 'Matched'), 5),
          omit = c('scale', 'dist_nearest04'),
          no.space = T,
          notes = 'Note: Scottish crime results only available from 2006 onwards',
          out = '../Results/ Inequality index model results.html',
          keep.stat = c('n')
          )

```

We can also fix the amenities that vary over time to monitor sensitivity. We can only compare fixed stats for years up
to 2015 / 2016. So fewer years leads to smaller N in the random effects model that use years observation.

```{r fixed}

fixed_crime.reg <- lmer(form1 %>% update(crime_fixed ~ .), reg.unmatch)

fixed_geo.reg <- lmer(form1 %>% update(geo_fixed ~ .), reg.unmatch)

fixed_pm25.reg <- lmer(form1 %>% update(pm25_fixed ~ .), reg.unmatch)

fixed_reg.list <- 
  list(
    crime.reg, fixed_crime.reg,
    geo.reg, fixed_geo.reg, 
    pm25.reg, fixed_pm25.reg
  )

stargazer(fixed_reg.list, type = 'text',
          title = 'Random effect model results',
          dep.var.caption = 'Inequality index (comparing varying vs fixed crime, access and pollution)',
          column.labels = c('Crime', 'Geo. access', 'Air pollution'),
          column.separate = c(2, 2, 2),
          dep.var.labels.include = F,
          
          covariate.labels = c('Year (2004 = 0)', 'Scotland', 'Year:Scotland'),
         add.lines =  list( c('Fixed values at 2004', rep(c('F', 'T'), 3)) ),
          omit = c('scale', 'dist_nearest04'),
          no.space = T,
          notes = 'Note: Scottish crime results uses 2006 SIMD values',
          out = '../Results/ Inequality index model results (fixed).html',
          keep.stat = c('n')
          )

```


