---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---


##  Note: Please run the source file first and any additional library
```{r setup, message = F }
knitr::opts_chunk$set(echo = T, message = F) #show code but no messages
```

```{r}
source('../UI paper 1 source.R')
library(MatchIt, quietly = T)
library(Matching, quietly = T)
library(stargazer, quietly = T)
library(lme4)
```

These are the graphing libraries and routines

```{r}
pkgs <- c('tidyverse', 'ggplot2', 'gridExtra')
lapply(pkgs, require, character.only = T)

hi.se <- function(x, t.val = 1.96 * sqrt(2) / 2){
  return(mean(x) + t.val * sd(x) / sqrt(length(x)))
}
lo.se <- function(x, t.val = 1.96 * sqrt(2) / 2){
  return(mean(x) - t.val * sd(x) / sqrt(length(x)))
}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

```



##  1) The sample

We first filter the data to non-isle and non-crossborder TTWAs. Then we only keep
TTWA with pops over 70k in 2004.

```{r}
ttwa.tab <- '../Results/Duncan index by TTWA.csv' %>% read.csv(stringsAsFactors = F)
crossborder.ttwa <- c('Hawick', 'Kelso & Jedburgh', 'Berwick', 'Carlisle') #cross national ttwas to be omitted
isles.ttwa <- c(
  'Eilean Siar',
  'Mull & Islay',
  'Campbeltown',
  'Penzance & Isles of Scilly',
  'Orkney Islands',
  'Shetland Islands',
  'Skye & Lochalsh',
  'Lochaber',
  'Irvine & Arran',
  'Isle of Wight')
#islands



ttwa.tab <- 
  ttwa.tab %>% 
  filter(!(ttwa %in% crossborder.ttwa)) %>% #not a cross border ttwa
  filter(!(ttwa %in% isles.ttwa)) %>% #nor isle
  ##  We need to take these out first since crossborder ttwa are duplicated
  group_by(ttwa) %>%
  mutate(pop04 = total.pop[year == '2004']) %>% #variable for pop in 2004
  ungroup %>%
  filter(pop04 > 70000) %>% ## over 70k in 2004
  filter(!is.na(ttwa)) ##finally no missign ttwa as this code is just due to our previous routine

```

We can restrict our data to 2004 to compare pre- and post- matching results.
```{r}
unmatch.tab <- 
  ttwa.tab %>% 
  filter(year == '2004')


unmatch.country <-
  unmatch.tab$country %>% table

unmatch.country %>% print

```

After omitting we get `r unmatch.country[1]` English TTWAs and `r unmatch.country[2]` Scottish TTWAs.



##  2) Matching using Genmatch

We have the full sample and now we further restrict to matched data. We match on 
2004 / time invariant. We can automatically try to ger the best match using genetic matching. Then we 
can compare the stae of our matches. Genetic matchign will use the propensity score 
in the matching by default (it generate own: i think from Sekhon's examples).

```{r, results= 'hide'}
## We need to hide all the print statement caused by genmatch
match.form <- I(country == 'Scotland') ~ total.pop + total.area + prop.inc

set.seed(123) ## genetic matchhing is a random process so we must set seed
match.res <- 
  matchit(match.form,
          data = unmatch.tab %>% dplyr::select(ttwa:prop.inc),
          method = 'genetic')
          
match.stats <-
  match.res %>% summary

```

Now we can try to summarise the matching results. 

```{r}
match.stats$nn

```

We have a total of `r match.stats$nn[1, 1]` English and `r match.stats$nn[1, 2]` Scottish ttwa. In our final matched sample, we have `r match.stats$nn[2, 1]` English TTWAs matched to`r match.stats$nn[2, 2]` Scottish ttwas. 

##  Comparing variance between indicies.

We note that the various indices are related to each other. 

```{r, message = T}
cor.df <- 
  unmatch.tab %>% filter(year == 2004) 

##  Making the correlation table 
cor.vars <- c('dist_nearest', 'suburb', 'di','live.in', 'crime', 'geo', 'pm25', 'work')
cor.table <- cor(cor.df %>% dplyr::select(cor.vars), 
                 method = 'spearman', 
                 use = 'pairwise.complete.obs') # only modest correlation

##  Formatting it for stargazer
cor.table[cor.table %>% upper.tri(diag = T)] <- NA
ind.nms <- c('RCI', 'Suburb. Index', 'Diss. Index', 'Housing', 'Crime', 'Geo. access', 'Air pollution', 'Work access')
rownames(cor.table) <- ind.nms
colnames(cor.table) <- ind.nms

stargazer(cor.table, 
          type = 'text', #anything else stargazer will output like code
          title = 'Spearman\'s rank correlation matrix between indicies (2004)',
          out = '../Results/ Index correlation table.html',
          notes = 'Note: Scottish crime data excluded from calculation')

```


##  Graphing trends over time

```{r}
rci.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = dist_nearest, colour = country)) + 
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean', size = 1) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Centralisation (RCI)') 
#rci.gg
#names(unmatched.tab)

sub.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = suburb, colour = country)) + 
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean', size = 1) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Surburbanisation')
#sub.gg

di.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = di, colour = country)) + 
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean', size = 1) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Uneveness (DI)')

## One big plot
mylegend<-g_legend(rci.gg + theme(legend.direction = 'horizontal'))

seg.grid <- grid.arrange(rci.gg + xlab('') + theme(legend.position="none"), 
                         sub.gg + xlab('') + theme(legend.position="none"), 
                         di.gg + xlab('')+ theme(legend.position="none"), 
                         ncol = 2,
                         #                    top = 'Changes in Segregation',
                         left = 'Index',
                         bottom = 'Year'
)

grid.arrange(mylegend, seg.grid, nrow=2,heights=c(1, 12),
             top = 'Change in segregation (Error bars for annual comparison)')


```

##  Inequality indices

```{r}

house.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = live.in, colour = country)) + 
  #  geom_point(position = 'jitter', alpha = 0.1) +
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               #               fun.args = (t.val = t.val = 1.96 * sqrt(2) / 2),
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean') +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Housing quality') 


crime.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = crime, colour = country)) + 
  #  geom_point(position = 'jitter', alpha = 0.1) +
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               #               fun.args = (t.val = 1.96 * sqrt(2) / 2),
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean') +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Crime')

##
geo.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = geo, colour = country)) + 
  #  geom_point(position = 'jitter', alpha = 0.1) +
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               #               fun.args = (t.val = 1.96 * sqrt(2) / 2),
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean') +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Geographical access')

air.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = pm25, colour = country)) + 
  #  geom_point(position = 'jitter', alpha = 0.1) +
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               #               fun.args = (t.val = 1.96 * sqrt(2) / 2),
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean', size = 1) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Exposure to particulates') 

work.gg <- 
  ggplot(data = ttwa.tab, mapping =aes(x = year, y = work, colour = country)) + 
  #  geom_point(position = 'jitter', alpha = 0.1) +
  stat_summary(geom = 'errorbar', 
               fun.ymax = hi.se, 
               fun.ymin = lo.se,
               #               fun.args = (t.val = t.val = 1.96 * sqrt(2) / 2),
               alpha = 0.5) +
  stat_summary(geom = 'line', fun.y = 'mean', size = 1) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  ylab('Access to employment') +
  xlab('')# 

##  Arrange ggplots

mylegend<-g_legend(house.gg + theme(legend.direction = 'horizontal'))


gfx <- grid.arrange(house.gg + xlab('') + theme(legend.position="none"), 
                    crime.gg + xlab('') + theme(legend.position="none"), 
                    geo.gg + xlab('')+ theme(legend.position="none"), 
                    air.gg + xlab('') + theme(legend.position="none"), 
                    work.gg + xlab('') + theme(legend.position="none"), 
                    #             top = 'Changes in inequality and centralisation',
                    left = 'Index',
                    bottom = 'Year'
)

grid.arrange(mylegend, gfx, nrow=2,heights=c(1, 12),
             top = 'Change in inequality (Error bars for annual comparison)')


```



##  Regression models

The data we are using for the models

```{r}
reg.unmatch <- ttwa.tab # just the whole data

##  For the matched data we basically need the TTWA and their weights
temp.match <- MatchIt::match.data(match.res)
head(temp.match)
temp.match <- temp.match %>% dplyr::select(ttwa, weights)
reg.match <- ttwa.tab %>% merge(temp.match, by = 'ttwa')

reg.match$year %>% table # okay so no missing
```



```{r}
##  The multilevel models 
form1 <- ~ I(year - 2004) + I(year - 2004)*country + (1|ttwa)
form2 <- form1 %>% update(~ . + scale(total.pop) + scale(total.area) + scale(prop.inc))
##Note:  form2 is form1 with more variables


# Segregation indicies regression ------------------------------------------------
##  
dist.reg <- lmer(form1 %>% update(dist_nearest ~ .), reg.unmatch)
dist.reg2 <- lmer(form2 %>% update(dist_nearest ~ .), reg.match)

sub.reg <- lmer(form1 %>% update(suburb ~ .), reg.unmatch)
sub.reg2 <- lmer(form2 %>% update(suburb ~ .), reg.match)

di.reg <- lmer(form1 %>% update(di ~ .), reg.unmatch)
di.reg2 <- lmer(form1 %>% update(di ~ .), reg.match)


## Save table
stargazer(dist.reg, dist.reg2,
          sub.reg, sub.reg2,
          di.reg, di.reg2,
          type = 'text',
          title = 'Model results for segregation',
          dep.var.caption = 'Segregation index',
          dep.var.labels = c('RCI', 'Suburbanisation', 'DI'),
          covariate.labels = c('Year (2004 = 0)', 'Scotland', 'Year:Scotland'),
          column.labels = rep(c('All', 'Matched'), 4),
          omit = c('scale'),
          no.space = T,
          out = '../Results/ Segregation index model results.html',
          keep.stat = c('n')
)

```

##  Regression by index

```{r}

in.reg <- lmer(form1 %>% update(live.in ~ .), reg.unmatch)
in.reg2 <- lmer(form2 %>% update(live.in ~ .), reg.match)

crime.reg <- lmer(form1 %>% update(crime ~ .), reg.unmatch)
crime.reg2 <- lmer(form2 %>% update(crime ~ .), reg.match)

geo.reg <- lmer(form1 %>% update(geo ~ .), reg.unmatch)
geo.reg2 <- lmer(form2 %>% update(geo ~ .), reg.match)

pm25.reg <- lmer(form1 %>% update(pm25 ~ .), reg.unmatch)
pm25.reg2 <- lmer(form2 %>% update(pm25 ~ .), reg.match)

work.reg <- lmer(form1 %>% update(work ~ .), reg.unmatch)
work.reg2 <- lmer(form2 %>% update(work ~ .), reg.match)

reg.list <- list(in.reg, in.reg2,
                 crime.reg, crime.reg2,
                 geo.reg, geo.reg2,
                 pm25.reg, pm25.reg2,
                 work.reg, work.reg2)

stargazer(reg.list, type = 'text',
          title = 'Random effect model results',
          dep.var.caption = 'Inequality index',
          dep.var.labels = c('Housing', 'Crime', 'Geo. access', 'Air pollution', 'Work access'),
          covariate.labels = c('Year (2004 = 0)', 'Scotland', 'Year:Scotland'),
          column.labels = rep(c('All', 'Matched'), 5),
          omit = c('scale'),
          no.space = T,
          notes = 'Note: Scottish crime results only available from 2006 onwards',
          out = '../Results/ Inequality index model results.html',
          keep.stat = c('n')
          )

```


This calls R inline `r 1`

